---
title: "Remote sensing of wildfire burn severity; Darkwoods Conservation Area"
author: "SMurphy"
date: "2020-08-29"
output:
  github_document:
    toc: TRUE
    toc_depth: 5
    df_print: tibble
#always_allow_html: yes
---

```{r setup, include=FALSE, message = FALSE, warning=FALSE, error=FALSE}
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
library(readxl)
library(sf)
library(kernlab)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(useful)
library(caret)
library(tibble)
library(klaR)
library(ModelMetrics)
library(DescTools)
library(knitr)
library(animation)
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE, 
  error=FALSE, comment = NA,tidy.opts=list(width.cutoff=60),tidy=TRUE)
# load the following variable for environment setup
#requirements = install.packages(c(
#  "readxl", "sf", "kernlab", "ggplot2", "dplyr", 
#  "RColorBrewer", "psych", "useful", "caret", 
#  "tibble", "klaR", "ModelMetrics", "DescTools", 
#  "MLmetrics", "knitr"), dependencies=TRUE)
#requirements
```

## 1. Model Designs

### 1.1 Import Data

Set seed for replication to `123`. Import excel data representing training data of burn severity classes from candidate spectral indices sampled in-field in composite burn plots.

```{r, warning=FALSE, message=FALSE}
set.seed(123) 
darkwoods_fire_plots_data <- read_excel("3.1.darkwoods_fire_ground_plots.xls")
```

### 1.2 Model Slope

```{r, warning=FALSE, message=FALSE}
ndbr_lm <- lm(CBI_total_2 ~ NDBR_2016, data = darkwoods_fire_plots_data)
ndbr_lm_resid <- resid(ndbr_lm)
plot(CBI_total_2 ~ NDBR_2016, data = darkwoods_fire_plots_data, 
     ylab = "Fire severity score of composite burn plots", xlab = "NDBR", col="blue")
abline(ndbr_lm, col = "red")
```

### 1.3 Model Training

Splitting data 70:30 between training and test subsets according to distribution of response variable: `CBI_total_2`.

```{r, warning=FALSE, message=FALSE}
training.samples <- createDataPartition(darkwoods_fire_plots_data$CBI_total_2, p=0.70, list = FALSE)
                                           train.data <- darkwoods_fire_plots_data[training.samples, ]
                                           test.data <- darkwoods_fire_plots_data[-training.samples, ]
```

Training regimes set by 10K-fold cross validation with 10, 5 and no repeats using `repeatedcv` and `cv` parameters.

```{r, warning=FALSE, message=FALSE}
model_training_10kfold_10repeat <- trainControl(method = "repeatedcv", 
                                      number = 10, repeats = 10,
                                      savePredictions = TRUE)


model_training_10kfold_3repeat <- trainControl(method = "repeatedcv", 
                                               number = 10, repeats = 3, 
                                               savePredictions = TRUE)


model_training_10kfold <- trainControl(method = "cv",
                                      number = 10, 
                                      savePredictions = TRUE) 
animation::cv.ani(k=10)
```

### 1.4 Model Tuning

Apply 10K-fold-X10 training regime to test prediction of NDBR between three candidate models including linear, non-linear support vector machine regressions, randomForest regression tree using `svmLinear` , `svmRadial` , `rf` kernels. Data preprocessing included using `center` and `scale` operations.

```{r, cache=TRUE, warning=FALSE, message=FALSE}
svm_ndbr_log_linear <- train(CBI_total_2 ~ NDBR_2016, 
                             data = train.data,
                             method = "svmLinear",
                             trControl = model_training_10kfold, 
                             preProcess = c("center","scale"), 
                             tuneGrid = expand.grid(C=seq(0,3, length = 10)),
                             metric="RMSE")

svm_ndbr_log_radial <- train(CBI_total_2 ~ NDBR_2016, 
                             data = train.data, 
                             method = "svmRadial",
                             trControl = model_training_10kfold, 
                             preProcess = c("center","scale"), 
                             tunelength=10)

rf_nndbr_1000trees = train(CBI_total_2 ~ NDBR_2016, 
                                    data = train.data, 
                                    method="rf", ntree=1000, 
                                    metric="RMSE",
                                    trControl=model_training_10kfold,
                                    importance = TRUE)

# Selected kernel
svm_ndbr_log_linear_full <- train(CBI_total_2 ~ NDBR_2016, 
                                  data = darkwoods_fire_plots_data,
                                  method = "svmLinear",
                                  trControl = model_training_10kfold, 
                                  preProcess = c("center","scale"), 
                                  tuneGrid = expand.grid(C=seq(0,3, length = 10)),
                                  metric="RMSE")

```

### 1.5 Model Validation

Compare performance of `svmLinear` model between training, test, and full dataset, using absolute and relative MAE, RMSE, and RMSE-ratio metrics, and `TheilU` statistic of model bias. Results tables x 2 below were taken from manuscript section 2.3 - both needing updates.

```{r, warning=FALSE, message=FALSE}
svm_ndbr_pred_train <- predict(svm_ndbr_log_linear, data = train.data)
svm_ndbr_pred_test <- predict(svm_ndbr_log_linear, data = test.data)
svm_ndbr_pred_full <- predict(svm_ndbr_log_linear_full, data = darkwoods_fire_plots_data)
svm_ndbr_pred_full_r2 = MLmetrics::R2_Score(svm_ndbr_pred_full, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_r2 = MLmetrics::R2_Score(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_r2 = MLmetrics::R2_Score(svm_ndbr_pred_test, test.data$CBI_total_2)
svm_ndbr_pred_full_mae <- mae(svm_ndbr_pred_full, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_mae <- mae(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_mae <- mae(svm_ndbr_pred_test, test.data$CBI_total_2)
svm_ndbr_pred_full_mae_rel <- (svm_ndbr_pred_full_mae/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_pred_train_mae_rel <- (svm_ndbr_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_ndbr_pred_test_mae_rel <- (svm_ndbr_pred_test_mae/mean(test.data$CBI_total_2))*100
svm_ndbr_pred_full_rmse <- rmse(svm_ndbr_pred_full, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_rmse <- rmse(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_rmse <- rmse(svm_ndbr_pred_test, test.data$CBI_total_2)
svm_ndbr_pred_full_rmse_rel <- (svm_ndbr_pred_full_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_pred_train_rmse_rel <- (svm_ndbr_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_ndbr_pred_test_rmse_rel <- (svm_ndbr_pred_test_rmse/mean(test.data$CBI_total_2))*100
svm_ndbr_pred_train_rmseRatio = svm_ndbr_pred_train_rmse/svm_ndbr_pred_test_rmse
svm_ndbr_pred_test_rmseRatio = svm_ndbr_pred_test_rmse/svm_ndbr_pred_train_rmse
svm_ndbr_full_Ubias_DescTools = DescTools::TheilU(darkwoods_fire_plots_data$CBI_total_2, svm_ndbr_pred_full, type = 2)

svm_ndbr_pred_full_r2
svm_ndbr_pred_train_r2
svm_ndbr_pred_test_r2
svm_ndbr_pred_full_mae
svm_ndbr_pred_train_mae 
svm_ndbr_pred_test_mae
svm_ndbr_pred_full_mae_rel
svm_ndbr_pred_train_mae_rel
svm_ndbr_pred_test_mae_rel
svm_ndbr_pred_full_rmse
svm_ndbr_pred_train_rmse
svm_ndbr_pred_test_rmse
svm_ndbr_pred_full_rmse_rel
svm_ndbr_pred_train_rmse_rel
svm_ndbr_pred_test_rmse_rel
svm_ndbr_pred_train_rmseRatio
svm_ndbr_pred_test_rmseRatio
svm_ndbr_full_Ubias_DescTools
svm_ndbr_log_linear_full$finalModel 
```

|           |             |          |                 |            |       |       |
|:---------:|:-----------:|:--------:|:---------------:|:----------:|:-----:|:-----:|
| **Model** |   **R2**    | **RMSE** | **RMSE^ratio^** | **TheilU** | **C** | **ε** |
|    NBR    | 0.879\*\*\* |   1.05   |      0.46       |   0.035    | 0.336 | 0.10  |
|  BAI2-SL  | 0.183\*\*\* |   2.61   |      1.09       |   0.307    | 1.105 | 0.10  |
|   MIRBI   | 0.551\*\*\* |   1.94   |      1.16       |   0.469    | 0.316 | 0.10  |
|    TVI    | 0.584\*\*\* |   1.88   |      1.16       |   0.432    | 0.474 | 0.10  |
|  TVI-SW   | 0.241\*\*\* |   2.54   |      1.10       |   0.343    | 0.158 | 0.10  |

## 2. Full Model List

```{r, warning=FALSE, message=FALSE}
#model 2 - NDBR-SW
svm_ndbr_sw_linear <- train(CBI_total_2 ~ NDBR_SW2_2, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 3 - BAI2
svm_bai2_linear <- train(CBI_total_2 ~ BAI2_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 4 - MIRBI
svm_mirbi_linear <- train(CBI_total_2 ~ MIRBI_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 5 - MSAV
svm_msav_linear <- train(CBI_total_2 ~ MSAV_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 6 - TVI
svm_tvi_linear <- train(CBI_total_2 ~ VIT2_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 7 - TVI-SW
svm_tvi_sw_linear <- train(CBI_total_2 ~ VIRSW2_201, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))
```

```{r, warning=FALSE, message=FALSE}
#model 2 - NDBRSW results
summary(lm(predict(svm_ndbr_sw_linear) ~ train.data$CBI_total_2))
svm_ndbr_sw_pred_train <- predict(svm_ndbr_sw_linear, data = train.data)
svm_ndbr_sw_pred_train_mae <- mae(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_mae 
svm_ndbr_sw_pred_train_mae_rel <- (svm_ndbr_sw_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_ndbr_sw_pred_train_mae_rel
svm_ndbr_sw_pred_train_rmse <- rmse(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_rmse
svm_ndbr_sw_pred_train_rmse_rel <- (svm_ndbr_sw_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_ndbr_sw_pred_train_rmse_rel
svm_ndbr_sw_pred_train_R2 <- R2(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_R2
TheilU(train.data$CBI_total_2, svm_ndbr_sw_pred_train, type = 2)
svm_ndbr_sw_train_Ubias = DescTools::TheilU(train.data$CBI_total_2, svm_ndbr_sw_pred_train, type = 2)

svm_ndbr_sw_train_Ubias <- ((svm_ndbr_sw_pred_train_mae)*68)/((svm_ndbr_sw_pred_train_mae)^2)
svm_ndbr_sw_train_Ubias

svm_ndbr_sw_pred_test <- predict(svm_ndbr_sw_linear, data = darkwoods_fire_plots_data)
svm_ndbr_sw_pred_test_rmse <- rmse(svm_ndbr_sw_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_sw_pred_test_rmse
svm_ndbr_sw_pred_test_rmse_rel <- (svm_ndbr_sw_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_sw_pred_test_rmse_rel

svm_ndbr_sw_pred_train_rmse/svm_ndbr_sw_pred_test_rmse
svm_ndbr_sw_pred_test_rmse/svm_ndbr_sw_pred_train_rmse

svm_ndbr_sw_linear
plot(svm_ndbr_sw_linear)
densityplot(svm_ndbr_sw_linear)
svm_ndbr_sw_linear$finalModel


#model 3 - BAI2 results
summary(lm(predict(svm_bai2_linear) ~ train.data$CBI_total_2))
svm_bai2_pred_train <- predict(svm_bai2_linear, data = train.data)
svm_bai2_pred_train_mae <- mae(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_mae 
svm_bai2_pred_train_mae_rel <- (svm_bai2_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_bai2_pred_train_mae_rel
svm_bai2_pred_train_rmse <- rmse(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_rmse
svm_bai2_pred_train_rmse_rel <- (svm_bai2_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_bai2_pred_train_rmse_rel
svm_bai2_pred_train_R2 <- R2(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_R2
TheilU(train.data$CBI_total_2, svm_bai2_pred_train, type = 2)
svm_bai2_train_Ubias <- ((svm_bai2_pred_train_mae)*68)/((svm_bai2_pred_train_mae)^2)
svm_bai2_train_Ubias

svm_bai2_pred_test <- predict(svm_bai2_linear, data = darkwoods_fire_plots_data)
svm_bai2_pred_test_rmse <- rmse(svm_bai2_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_bai2_pred_test_rmse
svm_bai2_pred_test_rmse_rel <- (svm_bai2_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_bai2_pred_test_rmse_rel

svm_bai2_pred_train_rmse/svm_bai2_pred_test_rmse
svm_bai2_pred_test_rmse/svm_bai2_pred_train_rmse

svm_bai2_linear
plot(svm_bai2_linear)
densityplot(svm_bai2_linear)
svm_bai2_linear$finalModel
plot(svm_bai2_linear)


#model 4 - MIRBI results
summary(lm(predict(svm_mirbi_linear) ~ train.data$CBI_total_2))
svm_mirbi_pred_train <- predict(svm_mirbi_linear, data = train.data)
svm_mirbi_pred_train_mae <- mae(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_mae 
svm_mirbi_pred_train_mae_rel <- (svm_mirbi_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_mirbi_pred_train_mae_rel
svm_mirbi_pred_train_rmse <- rmse(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_rmse
svm_mirbi_pred_train_rmse_rel <- (svm_mirbi_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_mirbi_pred_train_rmse_rel
svm_mirbi_pred_train_R2 <- R2(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_R2
TheilU(train.data$CBI_total_2, svm_mirbi_pred_train, type = 2)
svm_mirbi_train_Ubias <- ((svm_mirbi_pred_train_mae)*68)/((svm_mirbi_pred_train_mae)^2)
svm_mirbi_train_Ubias

svm_mirbi_pred_test <- predict(svm_mirbi_linear, data = darkwoods_fire_plots_data)
svm_mirbi_pred_test_rmse <- rmse(svm_mirbi_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_mirbi_pred_test_rmse
svm_mirbi_pred_test_rmse_rel <- (svm_mirbi_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_mirbi_pred_test_rmse_rel

svm_mirbi_pred_train_rmse/svm_mirbi_pred_test_rmse
svm_mirbi_pred_test_rmse/svm_mirbi_pred_train_rmse

svm_mirbi_linear
plot(svm_mirbi_linear)
densityplot(svm_mirbi_linear)
svm_mirbi_linear$finalModel
plot(svm_mirbi_linear)


#model 5 - MSAV results
summary(lm(predict(svm_msav_linear) ~ train.data$CBI_total_2))
svm_msav_pred_train <- predict(svm_msav_linear, data = train.data)
svm_msav_pred_train_mae <- mae(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_mae 
svm_msav_pred_train_mae_rel <- (svm_msav_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_msav_pred_train_mae_rel
svm_msav_pred_train_rmse <- rmse(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_rmse
svm_msav_pred_train_rmse_rel <- (svm_msav_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_msav_pred_train_rmse_rel
svm_msav_pred_train_R2 <- R2(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_R2
TheilU(train.data$CBI_total_2, svm_msav_pred_train, type = 2)
svm_msav_train_Ubias <- ((svm_msav_pred_train_mae)*68)/((svm_msav_pred_train_mae)^2)
svm_msav_train_Ubias

svm_msav_pred_test <- predict(svm_msav_linear, data = darkwoods_fire_plots_data)
svm_msav_pred_test_rmse <- rmse(svm_msav_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_msav_pred_test_rmse
svm_msav_pred_test_rmse_rel <- (svm_msav_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_msav_pred_test_rmse_rel

svm_msav_pred_train_rmse/svm_msav_pred_test_rmse
svm_msav_pred_test_rmse/svm_msav_pred_train_rmse

svm_msav_linear
plot(svm_msav_linear)
densityplot(svm_msav_linear)
svm_msav_linear$finalModel
plot(svm_msav_linear)


#model 6 - TVI results
summary(lm(predict(svm_tvi_linear) ~ train.data$CBI_total_2))
svm_tvi_pred_train <- predict(svm_tvi_linear, data = train.data)
svm_tvi_pred_train_mae <- mae(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_mae 
svm_tvi_pred_train_mae_rel <- (svm_tvi_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_tvi_pred_train_mae_rel
svm_tvi_pred_train_rmse <- rmse(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_rmse
svm_tvi_pred_train_rmse_rel <- (svm_tvi_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_tvi_pred_train_rmse_rel
svm_tvi_pred_train_R2 <- R2(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_R2
TheilU(train.data$CBI_total_2, svm_tvi_pred_train, type = 2)
svm_tvi_train_Ubias <- ((svm_tvi_pred_train_mae)*68)/((svm_tvi_pred_train_mae)^2)
svm_tvi_train_Ubias

svm_tvi_pred_test <- predict(svm_tvi_linear, data = darkwoods_fire_plots_data)
svm_tvi_pred_test_rmse <- rmse(svm_tvi_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_tvi_pred_test_rmse
svm_tvi_pred_test_rmse_rel <- (svm_tvi_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_tvi_pred_test_rmse_rel

svm_tvi_pred_train_rmse/svm_tvi_pred_test_rmse
svm_tvi_pred_test_rmse/svm_tvi_pred_train_rmse

svm_tvi_linear
plot(svm_tvi_linear)
densityplot(svm_tvi_linear)
svm_tvi_linear$finalModel
plot(svm_tvi_linear)


#model 7 - TVI-SW results
summary(lm(predict(svm_tvi_sw_linear) ~ train.data$CBI_total_2))
svm_tvi_sw_pred_train <- predict(svm_tvi_sw_linear, data = train.data)
svm_tvi_sw_pred_train_mae <- mae(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_mae 
svm_tvi_sw_pred_train_mae_rel <- (svm_tvi_sw_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_tvi_sw_pred_train_mae_rel
svm_tvi_sw_pred_train_rmse <- rmse(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_rmse
svm_tvi_sw_pred_train_rmse_rel <- (svm_tvi_sw_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_tvi_sw_pred_train_rmse_rel
svm_tvi_sw_pred_train_R2 <- R2(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_R2
TheilU(train.data$CBI_total_2, svm_tvi_sw_pred_train, type = 2)
svm_tvi_sw_train_Ubias <- ((svm_tvi_sw_pred_train_mae)*68)/((svm_tvi_sw_pred_train_mae)^2)
svm_tvi_sw_train_Ubias

svm_tvi_sw_pred_test <- predict(svm_tvi_sw_linear, data = darkwoods_fire_plots_data)
svm_tvi_sw_pred_test_rmse <- rmse(svm_tvi_sw_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_tvi_sw_pred_test_rmse
svm_tvi_sw_pred_test_rmse_rel <- (svm_tvi_sw_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_tvi_sw_pred_test_rmse_rel

svm_tvi_sw_pred_train_rmse/svm_tvi_sw_pred_test_rmse
svm_tvi_sw_pred_test_rmse/svm_tvi_sw_pred_train_rmse

svm_tvi_sw_linear
plot(svm_tvi_sw_linear)
densityplot(svm_tvi_sw_linear)
svm_tvi_sw_linear$finalModel
```
