---
title: "Remote sensing of wildfire burn severity; Darkwoods Conservation Area"
author: "SMurphy"
date: "2020-08-29"
output: github_document
---

```{r setup, include=FALSE}
#install.packages(c("readxl", "sf", "kernlab", "ggplot2", "dplyr", "RColorBrewer", "psych", "useful", "caret", "tibble", "klaR", "ModelMetrics", "DescTools", "MLmetrics"), dependencies=TRUE)
library(readxl)
library(sf)
library(kernlab)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(useful)
library(caret)
library(tibble)
library(klaR)
library(ModelMetrics)
library(DescTools)
#library(DescTools)
knitr::opts_chunk$set(echo = TRUE)
```

## Import & tidy

Set seed for replication to `123`. Import excel data representing training data of burn severity classes from candidate spectral indices sampled in-field in composite burn plots.

```{r}
set.seed(123) 
darkwoods_fire_plots_data <- read_excel("3.1.darkwoods_fire_plots.xls")
print(darkwoods_fire_plots_data)
```

# Model NDBR

```{r}
ndbr_lm <- lm(CBI_total_2 ~ NDBR_2016, data = darkwoods_fire_plots_data)
ndbr_lm_resid <- resid(ndbr_lm)
plot(CBI_total_2 ~ NDBR_2016, data = darkwoods_fire_plots_data, 
     ylab = "Fire severity score of composite burn plots", xlab = "NDBR", col="blue")
abline(ndbr_lm, col = "red")
```

## Train models

Splitting data 70:30 between training and test subsamples according to distribution of response variable: 'CBI_total_2'.

```{r}
training.samples <- createDataPartition(darkwoods_fire_plots_data$CBI_total_2, p=0.70, list = FALSE)
                                           train.data <- darkwoods_fire_plots_data[training.samples, ]
                                           test.data <- darkwoods_fire_plots_data[-training.samples, ]
```

A.  Model training regime fitted with 10K-fold cross validation by 10 repeats: `repeatedcv`. Use `savePredictions` function for visualization purposes.
B.  Model training regime fitted with 10K-fold cross validation by 3 repeats: `method = "repeatedcv", number = 10, repeats = 3`
C.  Model training regime fitted with simple 5K-fold cross validation: `method = "cv", number = 5`

```{r}
model_training_10kfold_10repeat <- trainControl(method = "repeatedcv", 
                                      number = 10, repeats = 10,
                                      savePredictions = TRUE)


model_training_10kfold_3repeat <- trainControl(method = "repeatedcv", 
                                               number = 10, repeats = 3, 
                                               savePredictions = TRUE)


model_training_5kfold <- trainControl(method = "cv",
                                      number = 5, 
                                      savePredictions = TRUE) 
```


## Tune models

Apply 10K-fold-X10 training regime to training data to test performance of support vector machine regressions with `svmRadial` and `svmLinear` kernels, and random forest regression tree. Models include preprocessing operations to `center` and `scale` distribution of datasets. Extent, patterns and grain of tuning grids are set differently with `tuneGrid` and `tuneLength` functions. 

1.  Model-1: NDBR prediction tested with linear SVM kernel and trained with 10K-fold-X10 regime.
2.  Model-2: NDBR prediction tested with nonlinear SVM kernel and trained with 10K-fold-X10 regime.
3.  Model-3: NDBR prediction tested with random forest regression tree of `1000` decision branches and performance bias `Rsquared`, and trained with 10K-fold-X10 regime. 

```{r}
svm_ndbr_log_linear <- train(CBI_total_2 ~ NDBR_2016, 
                             data = train.data,
                             method = "svmLinear",
                             trControl = model_training_10kfold_10repeat, 
                             preProcess = c("center","scale"), 
                             tuneGrid = expand.grid(C=seq(0,3, length = 10)),
                             metric="RMSE")

svm_ndbr_log_radial <- train(CBI_total_2 ~ NDBR_2016, 
                             data = train.data, 
                             method = "svmRadial",
                             trControl = model_training_10kfold_10repeat, 
                             preProcess = c("center","scale"), 
                             tunelength=10)

rf_nndbr_1000trees = train(CBI_total_2 ~ NDBR_2016, 
                                    data = train.data, 
                                    method="rf", ntree=1000, 
                                    metric="RMSE",
                                    trControl=model_training_10kfold_10repeat,
                                    importance = TRUE)

# Selected model
svm_ndbr_log_linear_full <- train(CBI_total_2 ~ NDBR_2016, 
                                  data = darkwoods_fire_plots_data,
                                  method = "svmLinear",
                                  trControl = model_training_10kfold_10repeat, 
                                  preProcess = c("center","scale"), 
                                  tuneGrid = expand.grid(C=seq(0,3, length = 10)),
                                  metric="RMSE")

svm_ndbr_log_linear_full$finalModel # Final parameters reported
```

## Test models

Compare model performances against test data measuring accuracy, precision, and model biases using absolute and relative MAE, RMSE, and RMSE-ratio metrics, as well as the `TheilU` function.

```{r}
#summary(lm(predict(svm_ndbr_log_linear) ~ train.data$CBI_total_2))
#summary(lm(predict(svm_ndbr_log_linear) ~ test.data$CBI_total_2))

svm_ndbr_pred_train <- predict(svm_ndbr_log_linear, data = train.data)
svm_ndbr_pred_test <- predict(svm_ndbr_log_linear, data = test.data)
svm_ndbr_pred_full <- predict(svm_ndbr_log_linear_full, data = darkwoods_fire_plots_data)

svm_ndbr_pred_full_r2 = MLmetrics::R2_Score(svm_ndbr_pred_full, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_r2 = MLmetrics::R2_Score(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_r2 = MLmetrics::R2_Score(svm_ndbr_pred_test, test.data$CBI_total_2)

svm_ndbr_pred_full_mae <- mae(svm_ndbr_pred_full, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_mae <- mae(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_mae <- mae(svm_ndbr_pred_test, test.data$CBI_total_2)
svm_ndbr_pred_full_mae_rel <- (svm_ndbr_pred_full_mae/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_pred_train_mae_rel <- (svm_ndbr_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_ndbr_pred_test_mae_rel <- (svm_ndbr_pred_test_mae/mean(test.data$CBI_total_2))*100

svm_ndbr_pred_full_rmse <- rmse(svm_ndbr_pred_train, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_pred_train_rmse <- rmse(svm_ndbr_pred_train, train.data$CBI_total_2)
svm_ndbr_pred_test_rmse <- rmse(svm_ndbr_pred_test, test.data$CBI_total_2)
svm_ndbr_pred_full_rmse_rel <- (svm_ndbr_pred_full_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_pred_train_rmse_rel <- (svm_ndbr_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_ndbr_pred_test_rmse_rel <- (svm_ndbr_pred_test_rmse/mean(test.data$CBI_total_2))*100

svm_ndbr_pred_train_rmseRatio = svm_ndbr_pred_train_rmse/svm_ndbr_pred_test_rmse
svm_ndbr_pred_test_rmseRatio = svm_ndbr_pred_test_rmse/svm_ndbr_pred_train_rmse
svm_ndbr_full_Ubias_DescTools = DescTools::TheilU(darkwoods_fire_plots_data$CBI_total_2, svm_ndbr_pred_full, type = 2)

svm_ndbr_pred_full_r2
svm_ndbr_pred_train_r2
svm_ndbr_pred_test_r2

svm_ndbr_pred_full_mae
svm_ndbr_pred_train_mae 
svm_ndbr_pred_test_mae

svm_ndbr_pred_full_mae_rel
svm_ndbr_pred_train_mae_rel
svm_ndbr_pred_test_mae_rel

svm_ndbr_pred_full_rmse
svm_ndbr_pred_train_rmse
svm_ndbr_pred_test_rmse

svm_ndbr_pred_full_rmse_rel
svm_ndbr_pred_train_rmse_rel
svm_ndbr_pred_test_rmse_rel

svm_ndbr_pred_train_rmseRatio
svm_ndbr_pred_test_rmseRatio
svm_ndbr_full_Ubias_DescTools
```

## Visualize models

```{r}
trellis.par.set(caretTheme())
plot(svm_ndbr_log_linear)
plot(svm_ndbr_log_radial)
#plot(rf_nndbr_1000trees)
densityplot(svm_ndbr_log_linear, pch = "|")
densityplot(svm_ndbr_log_radial, pch = "|")
densityplot(rf_nndbr_1000trees, pch = "|")
```

## Compare models

```{r}

#model 2 - NDBR-SW
svm_ndbr_sw_linear <- train(CBI_total_2 ~ NDBR_SW2_2, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 3 - BAI2
svm_bai2_linear <- train(CBI_total_2 ~ BAI2_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 4 - MIRBI
svm_mirbi_linear <- train(CBI_total_2 ~ MIRBI_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 5 - MSAV
svm_msav_linear <- train(CBI_total_2 ~ MSAV_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 6 - TVI
svm_tvi_linear <- train(CBI_total_2 ~ VIT2_2016, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))

#model 7 - TVI-SW
svm_tvi_sw_linear <- train(CBI_total_2 ~ VIRSW2_201, 
                            data = train.data,
                            method = "svmLinear",
                            trControl = model_training_10kfold, 
                            preProcess = c("center","scale"), 
                            tuneGrid = expand.grid(C=seq(0,3, length = 20)))


#model 2 - NDBRSW results
summary(lm(predict(svm_ndbr_sw_linear) ~ train.data$CBI_total_2))
svm_ndbr_sw_pred_train <- predict(svm_ndbr_sw_linear, data = train.data)
svm_ndbr_sw_pred_train_mae <- mae(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_mae 
svm_ndbr_sw_pred_train_mae_rel <- (svm_ndbr_sw_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_ndbr_sw_pred_train_mae_rel
svm_ndbr_sw_pred_train_rmse <- rmse(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_rmse
svm_ndbr_sw_pred_train_rmse_rel <- (svm_ndbr_sw_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_ndbr_sw_pred_train_rmse_rel
svm_ndbr_sw_pred_train_R2 <- R2(svm_ndbr_sw_pred_train, train.data$CBI_total_2)
svm_ndbr_sw_pred_train_R2
TheilU(train.data$CBI_total_2, svm_ndbr_sw_pred_train, type = 2)
svm_ndbr_sw_train_Ubias <- ((svm_ndbr_sw_pred_train_mae)*68)/((svm_ndbr_sw_pred_train_mae)^2)
svm_ndbr_sw_train_Ubias

svm_ndbr_sw_pred_test <- predict(svm_ndbr_sw_linear, data = darkwoods_fire_plots_data)
svm_ndbr_sw_pred_test_rmse <- rmse(svm_ndbr_sw_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_ndbr_sw_pred_test_rmse
svm_ndbr_sw_pred_test_rmse_rel <- (svm_ndbr_sw_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_ndbr_sw_pred_test_rmse_rel

svm_ndbr_sw_pred_train_rmse/svm_ndbr_sw_pred_test_rmse
svm_ndbr_sw_pred_test_rmse/svm_ndbr_sw_pred_train_rmse

svm_ndbr_sw_linear
plot(svm_ndbr_sw_linear)
densityplot(svm_ndbr_sw_linear)
svm_ndbr_sw_linear$finalModel
plot(svm_ndbr_sw_linear)


#model 3 - BAI2 results
summary(lm(predict(svm_bai2_linear) ~ train.data$CBI_total_2))
svm_bai2_pred_train <- predict(svm_bai2_linear, data = train.data)
svm_bai2_pred_train_mae <- mae(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_mae 
svm_bai2_pred_train_mae_rel <- (svm_bai2_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_bai2_pred_train_mae_rel
svm_bai2_pred_train_rmse <- rmse(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_rmse
svm_bai2_pred_train_rmse_rel <- (svm_bai2_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_bai2_pred_train_rmse_rel
svm_bai2_pred_train_R2 <- R2(svm_bai2_pred_train, train.data$CBI_total_2)
svm_bai2_pred_train_R2
TheilU(train.data$CBI_total_2, svm_bai2_pred_train, type = 2)
svm_bai2_train_Ubias <- ((svm_bai2_pred_train_mae)*68)/((svm_bai2_pred_train_mae)^2)
svm_bai2_train_Ubias

svm_bai2_pred_test <- predict(svm_bai2_linear, data = darkwoods_fire_plots_data)
svm_bai2_pred_test_rmse <- rmse(svm_bai2_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_bai2_pred_test_rmse
svm_bai2_pred_test_rmse_rel <- (svm_bai2_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_bai2_pred_test_rmse_rel

svm_bai2_pred_train_rmse/svm_bai2_pred_test_rmse
svm_bai2_pred_test_rmse/svm_bai2_pred_train_rmse

svm_bai2_linear
plot(svm_bai2_linear)
densityplot(svm_bai2_linear)
svm_bai2_linear$finalModel
plot(svm_bai2_linear)


#model 4 - MIRBI results
summary(lm(predict(svm_mirbi_linear) ~ train.data$CBI_total_2))
svm_mirbi_pred_train <- predict(svm_mirbi_linear, data = train.data)
svm_mirbi_pred_train_mae <- mae(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_mae 
svm_mirbi_pred_train_mae_rel <- (svm_mirbi_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_mirbi_pred_train_mae_rel
svm_mirbi_pred_train_rmse <- rmse(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_rmse
svm_mirbi_pred_train_rmse_rel <- (svm_mirbi_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_mirbi_pred_train_rmse_rel
svm_mirbi_pred_train_R2 <- R2(svm_mirbi_pred_train, train.data$CBI_total_2)
svm_mirbi_pred_train_R2
TheilU(train.data$CBI_total_2, svm_mirbi_pred_train, type = 2)
svm_mirbi_train_Ubias <- ((svm_mirbi_pred_train_mae)*68)/((svm_mirbi_pred_train_mae)^2)
svm_mirbi_train_Ubias

svm_mirbi_pred_test <- predict(svm_mirbi_linear, data = darkwoods_fire_plots_data)
svm_mirbi_pred_test_rmse <- rmse(svm_mirbi_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_mirbi_pred_test_rmse
svm_mirbi_pred_test_rmse_rel <- (svm_mirbi_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_mirbi_pred_test_rmse_rel

svm_mirbi_pred_train_rmse/svm_mirbi_pred_test_rmse
svm_mirbi_pred_test_rmse/svm_mirbi_pred_train_rmse

svm_mirbi_linear
plot(svm_mirbi_linear)
densityplot(svm_mirbi_linear)
svm_mirbi_linear$finalModel
plot(svm_mirbi_linear)


#model 5 - MSAV results
summary(lm(predict(svm_msav_linear) ~ train.data$CBI_total_2))
svm_msav_pred_train <- predict(svm_msav_linear, data = train.data)
svm_msav_pred_train_mae <- mae(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_mae 
svm_msav_pred_train_mae_rel <- (svm_msav_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_msav_pred_train_mae_rel
svm_msav_pred_train_rmse <- rmse(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_rmse
svm_msav_pred_train_rmse_rel <- (svm_msav_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_msav_pred_train_rmse_rel
svm_msav_pred_train_R2 <- R2(svm_msav_pred_train, train.data$CBI_total_2)
svm_msav_pred_train_R2
TheilU(train.data$CBI_total_2, svm_msav_pred_train, type = 2)
svm_msav_train_Ubias <- ((svm_msav_pred_train_mae)*68)/((svm_msav_pred_train_mae)^2)
svm_msav_train_Ubias

svm_msav_pred_test <- predict(svm_msav_linear, data = darkwoods_fire_plots_data)
svm_msav_pred_test_rmse <- rmse(svm_msav_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_msav_pred_test_rmse
svm_msav_pred_test_rmse_rel <- (svm_msav_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_msav_pred_test_rmse_rel

svm_msav_pred_train_rmse/svm_msav_pred_test_rmse
svm_msav_pred_test_rmse/svm_msav_pred_train_rmse

svm_msav_linear
plot(svm_msav_linear)
densityplot(svm_msav_linear)
svm_msav_linear$finalModel
plot(svm_msav_linear)


#model 6 - TVI results
summary(lm(predict(svm_tvi_linear) ~ train.data$CBI_total_2))
svm_tvi_pred_train <- predict(svm_tvi_linear, data = train.data)
svm_tvi_pred_train_mae <- mae(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_mae 
svm_tvi_pred_train_mae_rel <- (svm_tvi_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_tvi_pred_train_mae_rel
svm_tvi_pred_train_rmse <- rmse(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_rmse
svm_tvi_pred_train_rmse_rel <- (svm_tvi_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_tvi_pred_train_rmse_rel
svm_tvi_pred_train_R2 <- R2(svm_tvi_pred_train, train.data$CBI_total_2)
svm_tvi_pred_train_R2
TheilU(train.data$CBI_total_2, svm_tvi_pred_train, type = 2)
svm_tvi_train_Ubias <- ((svm_tvi_pred_train_mae)*68)/((svm_tvi_pred_train_mae)^2)
svm_tvi_train_Ubias

svm_tvi_pred_test <- predict(svm_tvi_linear, data = darkwoods_fire_plots_data)
svm_tvi_pred_test_rmse <- rmse(svm_tvi_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_tvi_pred_test_rmse
svm_tvi_pred_test_rmse_rel <- (svm_tvi_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_tvi_pred_test_rmse_rel

svm_tvi_pred_train_rmse/svm_tvi_pred_test_rmse
svm_tvi_pred_test_rmse/svm_tvi_pred_train_rmse

svm_tvi_linear
plot(svm_tvi_linear)
densityplot(svm_tvi_linear)
svm_tvi_linear$finalModel
plot(svm_tvi_linear)


#model 7 - TVI-SW results
summary(lm(predict(svm_tvi_sw_linear) ~ train.data$CBI_total_2))
svm_tvi_sw_pred_train <- predict(svm_tvi_sw_linear, data = train.data)
svm_tvi_sw_pred_train_mae <- mae(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_mae 
svm_tvi_sw_pred_train_mae_rel <- (svm_tvi_sw_pred_train_mae/mean(train.data$CBI_total_2))*100
svm_tvi_sw_pred_train_mae_rel
svm_tvi_sw_pred_train_rmse <- rmse(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_rmse
svm_tvi_sw_pred_train_rmse_rel <- (svm_tvi_sw_pred_train_rmse/mean(train.data$CBI_total_2))*100
svm_tvi_sw_pred_train_rmse_rel
svm_tvi_sw_pred_train_R2 <- R2(svm_tvi_sw_pred_train, train.data$CBI_total_2)
svm_tvi_sw_pred_train_R2
TheilU(train.data$CBI_total_2, svm_tvi_sw_pred_train, type = 2)
svm_tvi_sw_train_Ubias <- ((svm_tvi_sw_pred_train_mae)*68)/((svm_tvi_sw_pred_train_mae)^2)
svm_tvi_sw_train_Ubias

svm_tvi_sw_pred_test <- predict(svm_tvi_sw_linear, data = darkwoods_fire_plots_data)
svm_tvi_sw_pred_test_rmse <- rmse(svm_tvi_sw_pred_test, darkwoods_fire_plots_data$CBI_total_2)
svm_tvi_sw_pred_test_rmse
svm_tvi_sw_pred_test_rmse_rel <- (svm_tvi_sw_pred_test_rmse/mean(darkwoods_fire_plots_data$CBI_total_2))*100
svm_tvi_sw_pred_test_rmse_rel

svm_tvi_sw_pred_train_rmse/svm_tvi_sw_pred_test_rmse
svm_tvi_sw_pred_test_rmse/svm_tvi_sw_pred_train_rmse

svm_tvi_sw_linear
plot(svm_tvi_sw_linear)
densityplot(svm_tvi_sw_linear)
svm_tvi_sw_linear$finalModel

plot(svm_tvi_sw_linear)
```

